Entropy stable skew-symmetric discontinuous Galerkin formulations on non-conforming meshes

### 2018-07-19 docs/ESDG_GQ/ext_decoupled_sbp.tex
updating ESDG GQ paper

%\documentclass[preprint,10pt]{elsarticle}
%\documentclass[preprint,10pt]{article}
\documentclass[review]{siamart0216}
%\documentclass{siamart0216}

@@

\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{amsthm}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}
%\theoremstyle{lemma}
%\newtheorem{lemma}{Lemma}
%\theoremstyle{corollary}
%\newtheorem{corollary}{Corollary}
%\theoremstyle{theorem}
%\newtheorem{theorem}{Theorem}
\newtheorem*{remark}{Remark}
\theoremstyle{assumption}
\newtheorem{assumption}{Assumption}

@@

\renewcommand{\hat}[1]{\widehat{#1}}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

@@

\date{}
\author{Jesse Chan}
\title{Skew-symmetric entropy stable discontinuous Galerkin formulations with application to hybrid meshes }

\graphicspath{{./figs/}}

@@

%\begin{frontmatter}
%\title{Entropy stable skew-symmetric discontinuous Galerkin formulations with applications to hybrid and non-conforming meshes. }
%
%\author[rice]{Jesse Chan\corref{cor1}}
%\ead{Jesse.Chan@caam.rice.edu}
%\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
%
%\begin{abstract}
%Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
%\end{abstract}
%\end{frontmatter}

@@

High order methods for the simulation of time-dependent compressible flow have the potential to achieve higher levels of accuracy at lower costs compared to current low order schemes \cite{wang2013high}.  In addition to superior accuracy and efficiency on modern computing architectures, the low numerical dispersion and dissipation of high order methods \cite{ainsworth2004dispersive} enables the accurate propagation of waves over long distances and time scales.  The same properties also make high order methods attractive for the resolution of unsteady phenomena such as vorticular and turbulent flows, which are sensitive to numerical dissipation \cite{visbal1999high, wang2013high}.  However, while high order methods are often provably stable for wave problems, high order schemes for nonlinear conservation laws have consistently been hampered by problems of instability.

When applied to nonlinear conservation laws, high order methods can experience artificial growth and blow-up near under-resolved features such as shocks, turbulence, or boundary layers.  In practice, the application of high order methods to practical problems requires shock capturing and stabilization techniques (such as artificial viscosity) or solution regularization (such as filtering or limiting) to prevent solution blow-up.  The resulting schemes for nonlinear conservation laws walk a fine line between stability, robustness, and accuracy.  Aggressive stabilization or regularization can result in the loss of high order accuracy, while too little can result in instability \cite{wang2013high}.  Moreover, it can be difficult to determine robust expressions for stabilization paramaters, as parameters which work for one simulation can fail when applied to a different physical regime or discretization setting.

These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.  Proofs of continuous entropy inequalities rely on the chain rule.  In contrast, these discrete entropy inequalities account for the loss of the chain rule due to effects such as quadrature errors, which are incurred when applying polynomially exact quadrature rules to nonlinear and rational integrands within the DG formulation.  These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely}.  These methods have also been extended to staggered grid methods \cite{parsani2016entropy} and non-conforming meshes \cite{friedrich2017entropy}.  Entropy stable boundary conditions have also been determined for the compressible Euler and Navier-Stokes equations \cite{parsani2015entropy, svard2018entropy}.

\section{Entropy stability for systems of nonlinear conservation laws}

We begin by reviewing the dissipation of entropy for a $d$-dimensional system of nonlinear conservation laws on a domain $\Omega$
\begin{equation}
\pd{\bm{u}}{t}  + \sum_{j=1}^d\pd{\bm{f}_j(\bm{u})}{x_j} = 0, \qquad \bm{u}\in \mathbb{R}^n, \qquad \bm{f}:\mathbb{R}^n\rightarrow\mathbb{R}^n,
\label{eq:nonlineqs}
\end{equation}
where $\bm{u}$ are the conservative variables and $\bm{f}(\bm{u})$ is a vector-valued nonlinear flux function.  We are interested in nonlinear conservation laws for which a convex entropy function $U(\bm{u})$ exists.  For such systems, the  \emph{entropy variables} are an invertible mapping $\bm{v}(\bm{u}):\mathbb{R}^n\rightarrow \mathbb{R}^n$ defined as the derivative of the entropy function with respect to the conservative variables 
\begin{align}
\bm{v}(\bm{u}) = \pd{U}{\bm{u}}.%, \qquad  \text{ invertible}.%\bm{u}\LRp{\bm{v}(\bm{u})} = \bm{u}.
\label{eq:entropyvarsmap}
\end{align}
Several widely used equations in fluid modeling (Burgers, shallow water, compressible Euler and Navier-Stokes equations) yield convex entropy functions $U(\bm{u})$ \cite{hughes1986new, chen2017entropy}.  Let $\partial \Omega$ be the boundary of $\Omega$ with outward unit normal $\bm{n}$.  By multiplying the equation (\ref{eq:nonlineqs}) with $\bm{v}(\bm{u})^T$, the solutions $\bm{u}$ of (\ref{eq:nonlineqs}) can be shown to satisfy an entropy inequality
\begin{equation}
\int_{\Omega}\pd{U(\bm{u})}{t}\diff{x} + \int_{\partial \Omega} \sum_{j=1}^d \LRp{\bm{v}(\bm{u})^T\bm{f}_j(\bm{u}) - \psi_j\LRp{\bm{v}(\bm{u})}}n_j \diff{x} \leq 0, \qquad \bm{n} = \LRp{n_1,\ldots,n_d}, 
\label{eq:entropyineq}
\end{equation}
where $\psi_j(\bm{u})$ is a function referred to as the entropy potential.

The proof of (\ref{eq:entropyineq}) requires the use of the chain rule \cite{mock1980systems, harten1983symmetric, dafermos2005compensated}.  The instability-in-practice of high order schemes for (\ref{eq:nonlineqs}) can be attributed in part to the fact that the discrete form of the equations do not satisfy the chain rule, and thus do not satisfy (\ref{eq:entropyineq}).  As a result, discretizations of (\ref{eq:nonlineqs}) do not typically possess an underlying statement of stability.  For low order schemes, this can be offset in practice by the inherent numerical dissipation.  However, because high order discretizations possess low numerical dissipation, the lack of an underlying discrete stability has been conflated with the idea that high order methods are inherently less stable than low order methods.

\section{Approximation spaces}

We define an approximation space using degree $N$ polynomials on the reference element.  In this work, we consider either simplicial reference elements (triangles and tetrahedra) or tensor product reference elements (quadrilaterals and hexahedra).  The natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.

On a $d$-dimensional reference simplex, the natural polynomial space are total degree $N$ polynomials 
\[
P^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq \sum_{k=1}^d i_k \leq N}.
\]
In contrast, the natural polynomial space on a $d$-dimensional tensor product element is 
\[
Q^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq i_k \leq N, \quad k = 1,\ldots, d}.
\]
We take $V^N$ to denote the natural approximation space on a given element.  Furthermore, we denote the dimension of $V^N$  as $N_p = {\rm dim}\LRp{V^N\LRp{\widehat{D}}}$.

We also define trace spaces for each reference element.  Let $\hat{f}$ be a face of the reference element $\hat{D}$.  The trace space $V^N \LRp{\hat{f}}$ is defined as the restrictions of functions in $V^N\LRp{\hat{D}}$ to $\hat{f}$
\[
V^N \LRp{\hat{f}} = \LRc{ \left.u\right|_{\hat{f}}, \quad u \in V^N\LRp{\hat{D}}}, \qquad \hat{f}\in \partial\hat{D}.
\]
For example, the trace space for degree $N$ polynomials on a quadrilateral face $\hat{f}$ of the bi-unit hexahedral element $[-1,1]^3$ is $Q^N\LRp{\hat{f}}$, while the trace space for degree $N$ polynomials on a triangular face $f$ of the tetrahedron is $P^N\LRp{\hat{f}}$.  We denote the dimension of the trace space as ${\rm dim}\LRp{V^N\LRp{{\hat{f}}}} = N^f_p$.

We similarly define the trace space for the surface $\partial \hat{D}$ of $\hat{D}$
\[
V^N \LRp{\partial \hat{D}} = \LRc{ \left.u\right|_{\partial \hat{D}}, \quad u \in V^N\LRp{\hat{D}}}.
\]

\section{$L^2$ projection and lifting matrices and a decoupled SBP operator}

The schemes in \cite{chan2017discretely} begin on the reference element $\hat{D}$ by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$.  It is assumed that volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ are provided for $\hat{D}$ and the surface $\partial \hat{D}$ of $\hat{D}$, respectively.  We make the following assumption on the volume quadrature rule:
\begin{assumption}
The volume quadrature rule integrates exactly all polynomials up to degree $(2N-1)$.
\label{ass:quad}
\end{assumption}
In addition to being related to the positive definiteness of the mass matrix, Assumption~\ref{ass:quad} ensures that, for any $u,v \in V^N$, the following integral is exactly approximated using quadrature
\[
\int_{\hat{D}} \pd{u(\hat{\bm{x}})}{\hat{x}_i} v(\hat{\bm{x}})\diff{\hat{\bm{x}}} = \sum_{k=1}^{N_q} w_k \pd{u(\hat{\bm{x}}_k)}{\hat{x}_i} v(\hat{\bm{x}}_k).  
\]
In other words, evaluating the integral of the derivative of $u$ and $v$ using quadrature is exact because the integrand is a polynomial of degree $(2N-1)$.

\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}

Let $\bm{V}_q,\bm{V}_f$ denote interpolation matrices, and let $\bm{D}^i$ be the differentiation matrix with respect to the $i$th coordinate such that
\begin{gather}
\LRp{\bm{V}_q}_{ij} = \phi_j(\bm{x}_i), \qquad \LRp{\bm{V}_f}_{ij} = \phi_j(\bm{x}^f_i), \qquad \pd{\phi_j(\bm{x})}{x_i} = \sum_{k=1}^{N_p} \LRp{\bm{D}^i_{jk}} \phi_k(\bm{x}).
\end{gather}
The interpolation matrices $\bm{V}_q,\bm{V}_f$ map basis coefficients to evaluations at volume and surface quadrature points respectively, while the differentiation matrix ${\bm{D}}_i$ maps basis coefficients of a function to the basis coefficients of its derivative with respect to $x_k$.  The interpolation matrices are used to assemble the mass matrix $\bm{M}$, the quadrature-based projection matrix $\bm{P}_q$, and lifting matrix $\bm{L}_f$
\begin{gather}
\bm{M} = \bm{V}_q^T\bm{W}\bm{V}_q, \qquad \bm{P}_q = \bm{M}^{-1}\bm{V}_q^T\bm{W}, \qquad \bm{L}_f = \bm{M}^{-1}\bm{V}_f^T\bm{W}_f,
\end{gather}
where $\bm{W}, \bm{W}_f$ are diagonal matrices of volume and surface quadrature weights, respectively.  The matrix $\bm{P}_q$ is a quadrature-based discretization of the $L^2$ projection operator $\Pi_N$ onto degree $N$ polynomials.

Interpolation, differentiation, and $L^2$ projection matrices can be combined to construct finite difference operators.  For example, the matrix $\bm{D}_q = \bm{V}_q\bm{D}^i\bm{P}_q$ maps function values at quadrature points to approximate values of the derivative at quadrature points.  By choosing specific quadrature rules, $\bm{D}_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.  However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, the PI introduced in \cite{chan2017discretely} a new ``decoupled'' summation by parts matrix which builds interface terms directly into the approximation of the derivative.

Let $\bm{n}$ denote the outward unit normal $\bm{n}$ on $\hat{D}$, and let $\bm{n}_i$ denote the vector containing values of the $i$th component at all surface quadrature points.  The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
\begin{align}
\bm{D}^i_N  &= \LRs{
\begin{array}{cc}
\bm{D}^i_q - \frac{1}{2}\bm{V}_q \bm{L}_f {\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\bm{V}_q\bm{L}_f{\rm diag}({\bm{n}}_i)\\
-\frac{1}{2}{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}{\rm diag}({\bm{n}}_i)
\end{array}}.  
\label{eq:DN}
\end{align}
When combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
\begin{gather}
f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
\label{eq:dnapprox}
\end{gather}
The approximation can also be interpreted as solving a variational problem.  Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.  Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that 
\[
\int_{\hat{D}} u(\bm{x})v(\bm{x})  = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}n_i, \qquad \forall v\in V^N\LRp{\hat{D}}.
\]
If the surface quadrature rule is exact for polynomials of degree $2N$, then the decoupled SBP operator $\bm{D}^i_N$ also satisfies a summation by parts property.  
\begin{gather}
\bm{Q}^i_N = \bm{W}_N \bm{D}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:dsbp}\\ \bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\nonumber
\end{gather}
By noting that $\bm{W}\bm{V}_q \bm{L}_q = \bm{P}_q^T\bm{V}_f^T\bm{W}_f$, the entries of $\bm{Q}^i_N$ can be re-written as
\begin{align}
\bm{Q}_i &= \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q,\\
\bm{Q}^i_N  &= \LRs{
\begin{array}{cc}
\bm{Q}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\\
-\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)
\end{array}}.
\label{eq:QN}
\end{align}
from which the SBP property (\ref{eq:dsbp}) is more apparent.

%\begin{lemma}[Lemma 1 in \cite{chan2017discretely}]
%Let $\bm{D}^i_q = \bm{V}_q\bm{D}^i\bm{P}_q$.  Then, for volume quadratures which are exact for degree $2N-1$ polynomials, $\bm{D}^i_q$ satisfies the SBP property with respect to $\bm{W}$ 
%\[
%\bm{W}\bm{D}^i_q = \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f\diag{\hat{\bm{n}}} \bm{V}_f\bm{P}_q - \LRp{\bm{W}\bm{D}^i_q}^T.  
%\]
%\label{lemma:gsbp}
%\end{lemma}

\section{Skew-symmetric entropy conservative formulations}

\subsection{Loss of the SBP property for hybrid and non-conforming meshes}

While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.  This difficulty is a result of the choices of surface quadrature which naturally arise for these settings.

\begin{figure}
\centering
\begingroup
\captionsetup[subfigure]{width=.425\textwidth}
\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
\endgroup
\hspace{2em}
\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
\caption{Examples of interface couplings which do not result in an SBP property and are not entropy stable.  On the left, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.  On the right, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral.  }
\label{fig:hybrid}
\end{figure}

\begin{figure}
\centering
\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
\hspace{2em}
\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
\label{fig:hybrid3d}
\end{figure}

\begin{figure}[!h]
\centering
\begingroup
\captionsetup[subfigure]{width=.45\textwidth}
\subfloat[Entropy stable inter-element coupling in \cite{friedrich2017entropy} for a non-conforming interface.]{\raisebox{0em}{\includegraphics[width=.45\textwidth]{figs/nonconSBP.png}}\label{subfig:noncon2}}
\hspace{2em}
\subfloat[Entropy stable inter-element coupling in this work for a non-conforming interface.  The dotted black lines denote communication between neighboring elements.]{\raisebox{-0em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
\endgroup
%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.  Each dashed red line indicates a dyadic flux computation required between two nodes. } %The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces.  The new approach results in a simpler communication pattern between elements. }
\label{fig:noncon}
\end{figure}

\subsection{A variational SBP property}

It can be shown that the SBP property holds for $\bm{D}^i_N$ if the surface quadrature is exact for polynomials of degree $2N$ \cite{chan2017discretely}.  The SBP property is lost if the surface quadrature is only exact for polynomial integrands of degree less than $2N$.  However, one can still show that a variational version of the SBP property is satisfied.

\begin{lemma}
Let $v\in V^N$, and let the surface trace $\LRl{v}_{\partial \hat{D}} \in V^M\LRp{\partial \hat{D}}$ for some $M \leq N$. If the surface quadrature is exact for polynomials of degree $N+M$,  then for an arbitrary vector $\bm{u}_q$, 
\[
\bm{v}_q^T\bm{Q}_i \bm{u}_q = \bm{v}_q^T\LRp{ \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f \diag{{\bm{n}}_i}\bm{V}_f\bm{P}_q - \bm{Q}_i^T}\bm{u}_q
\]
where $\bm{v}$ denotes the vector of values of $v$ of at both volume and surface quadrature points. 
\label{lemma:vsbp}
\end{lemma}
\begin{proof}
Recall that $\bm{Q}_i = \bm{W} \bm{V}_q \bm{D}^i\bm{P}_q$.  Since the volume quadrature is exact for degree $2N-1$ polynomials, we have that
\begin{align*}
\bm{v}_q^T\bm{Q}_i \bm{u}_q &= \bm{v}_q^T\bm{W} \bm{D}^i \bm{P}_q \bm{u}_q = \int_{\hat{D}} \pd{\Pi_N u}{\hat{x}_i} v = \int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i - \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i}.
\end{align*}
Since the volume integrand $\LRp{\Pi_N u} \pd{v}{\hat{x}_i} \in V^{2N-1}$, it is exactly integrated using the volume quadrature.  Additionally, since $\Pi_N u \in V^N$ and $v\in V^M$ on the surface of $\hat{D}$, the surface quadrature is exact for the surface integral and 
\begin{align*}
&\int_{\partial \hat{D}} (\Pi_N u) v \hat{n}_i -   \int_{\hat{D}} \LRp{\Pi_N u} \pd{v}{\hat{x}_i} =\\
 &\bm{v}_f^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \LRp{\bm{V}_q\bm{D}^i\bm{P}_q\bm{v}_q}^T\bm{W} \bm{V}_q\bm{P}_q\bm{u}_q\\
=& \bm{v}_q^T \LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q\bm{u}_q - \bm{v}_q^T \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q\bm{u}_q,
\end{align*}
where we have used that, since $v\in V^N$, $\bm{v}_f = \bm{V}_f\bm{P}_q\bm{v}_q$.  The proof is completed by noting that 
\[
\bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T \bm{W} \bm{V}_q\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{M}\bm{P}_q = \bm{P}_q^T\LRp{\bm{D}^i}^T \bm{V}_q^T\bm{W} = \bm{Q}_i^T.
\]
\end{proof}

The variational SBP property also extends to the decoupled SBP operator.  This property, along with the exact differentiation of constants, is necessary for the proof of entropy stability.  
\begin{lemma}
Let $\bm{D}^i_N$ be a decoupled SBP operator on the reference element $\hat{D}$, and let the surface quadrature be exact for polynomials of degree $N+M$ for some $M \leq N$.  Suppose $v\in V^N$ and that $\LRl{v}_{\partial \hat{D}} \in P^M\LRp{\partial \hat{D}}$, Then, 
\[
\bm{v}^T\bm{Q}^i_N\bm{u} = \bm{v}^T\LRp{\bm{B}^i_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.%, \qquad \bm{Q}^i_N \bm{1} = \bm{0}.
\]
where $\bm{v}$ denotes the values of $v$ of at both volume and surface quadrature points.  
\label{lemma:vdsbp}
\end{lemma}
\begin{proof}
For convenience, let $\bm{u}_q, \bm{u}_f$ denote evaluations of $u$ at volume and surface points, such that 
\[
\bm{u} = \begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}, \qquad \bm{v} = \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}.  
\]
The proof of the variational summation by parts property uses the definition of $\bm{Q}^i_N$ (\ref{eq:QN}), 
\begin{align*}
\bm{v}^T\bm{Q}^i_N\bm{u} &= \bm{v}_q^T\bm{Q}_i \bm{u}_q - \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
& - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f.
\end{align*}
Applying Lemma~\ref{lemma:vsbp} then yields
\begin{align*}
\bm{v}^T\bm{Q}^i_N\bm{u} =& -\bm{v}_q^T\bm{Q}^T_i \bm{u}_q + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q\bm{v}_q}^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
&\qquad - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \LRp{\bm{V}_f\bm{P}_q\bm{u}_q} - \frac{1}{2}\bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f + \bm{v}_f^T\bm{W}_f\hat{\bm{n}} \bm{u}_f\\
=& \begin{pmatrix} \bm{v}_q\\ \bm{v}_f\end{pmatrix}^T 
\left(\begin{pmatrix}
\bm{0}& \\
& \bm{W}_f\diag{\hat{\bm{n}}}
\end{pmatrix}\right.\\
&+
\left.\begin{pmatrix}
-\bm{Q}_i^T + \frac{1}{2}\LRp{\bm{V}_f\bm{P}_q}^T\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2} \LRp{\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q}^T\\
\frac{1}{2}\bm{W}_f\hat{\bm{n}} \bm{V}_f\bm{P}_q & -\frac{1}{2}\bm{W}_f\hat{\bm{n}}
\end{pmatrix}  \right)
\begin{pmatrix} \bm{u}_q\\ \bm{u}_f\end{pmatrix}\\
=& \bm{v}^T\LRp{\bm{B}_N - \LRp{\bm{Q}^i_N}^T}\bm{u}.
\end{align*}
\end{proof}
Using Lemma~\ref{lemma:vdsbp}, we will determine the accuracy of the surface quadrature necessary to ensure the proof of semi-discrete entropy stability is valid.  This minimal degree of surface quadrature accuracy will depend on the nature of the reference-to-physical mapping (e.g.\ affine vs curved elements).

>>>>

%\documentclass[preprint,10pt]{elsarticle}
%\documentclass[preprint,10pt]{amsart}
\documentclass[review]{siamart0216}
%\documentclass{siamart0216}

@@

\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{amsthm}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}
%\theoremstyle{lemma}
%\newtheorem{lemma}{Lemma}
%\theoremstyle{corollary}
%\newtheorem{corollary}{Corollary}
%\newtheorem*{remark}{Remark}
%\theoremstyle{theorem}
%\newtheorem{theorem}{Theorem}
%\theoremstyle{assumption}
%\newtheorem{assumption}{Assumption}

@@

\renewcommand{\tilde}{\widetilde}
\renewcommand{\hat}{\widehat}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

@@

%\date{}
\author{Jesse Chan, David C.\ Del Rey Fernandez}
\title{Entropy stable skew-symmetric discontinuous Galerkin formulations on non-conforming meshes}
\graphicspath{{./figs/}}

@@

%\begin{frontmatter}
%\title{Skew-symmetric formulations for entropy stable discontinuous Galerkin methods on non-conforming meshes}
%
%\author[rice]{Jesse Chan\corref{cor1}}
%\ead{Jesse.Chan@caam.rice.edu}
%\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
%
%\begin{abstract}
%Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
%\end{abstract}
%\end{frontmatter}

@@

\section{Entropy stable Gauss-Legendre collocation methods}

\note{Idea: decoupled SBP operators allow for the construction of entropy stable schemes by chaining together different sets of nodes through extrapolation (interpolation) operators.  }

\section{Gauss-Legendre collocation on non-conforming meshes}

On conforming meshes, it is most efficient to utilize both Gauss quadrature for volume integrals and Gauss quadrature for face or surface integrals.  For solutions represented in terms of their values at tensor product volume Gauss nodes, extrapolation to face Gauss nodes can be done in an efficient line-by-line manner using one-dimensional interpolation matrices.

For non-conforming meshes, it can be advantageous to use composite Gauss quadratures on non-conforming interfaces.  \note{Reference JK and LCW's paper on full-side vs split side mortars.}  However, interpolating the solution at volume Gauss nodes to split-side Gauss nodes is no longer a one-dimensional operation.

\begin{figure}
\centering
\begingroup
\captionsetup{width=.25\linewidth}
\subfloat[Conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss1.png}}
\hspace{4em}
\subfloat[Non-conforming faces]{\includegraphics[width=.25\textwidth]{figs/gauss2.png}}
\endgroup
\caption{Examples of surface quadratures for conforming and non-conforming interfaces for Gauss collocation methods.  }
\label{fig:gqcon_noncon}
\end{figure}

It is possible to efficiently interpolate from volume Gauss nodes to split-side Gauss nodes in two steps.  We first interpolate to full-side Gauss nodes on the face, then interpolate from full-side  to split-side Gauss nodes.  Each of these two operations can be done using only one-dimensional interpolation matrices, implying that the interpolation matrix from volume to split-side Gauss nodes can be decomposed into 
\[
\bm{V}_{\rm vol-to-split} = \bm{V}_{\rm vol-to-full} \bm{V}_{\rm full-to-split}
\]
where each of the matrices in the decomposition possess either a sparse or Kronecker product structure.

ESDG schemes based on decoupled SBP operators introduced in \cite{chan2017discretely} assume a single interpolation matrix which extrapolates from volume to surface nodes.  These schemes require dyadic flux evaluations between all volume and surface nodes connected by this interpolation matrix.  For non-conforming interfaces with split-side Gauss nodes on the face, this necessitates dyadic flux computations between all volume nodes and all split-side Gauss nodes.

We show in this section how to construct an entropy stable Gauss collocation scheme based on two layers of surface nodes: full-side and split-side nodes.  These surface nodes will be loosely through a skew-symmetric coupling term involving polynomial interpolation matrices between full-side and split-side nodes.  This scheme requires dyadic flux computations between volume and full-side surface nodes, and between full-side and split-side surface nodes.  Each set of dyadic flux computations can be reduced to

Let ${\bm{V}_f}\tilde{P}_f$ denote the matrix which projects to the trace space using split-side quadrature and evaluates at full-side quadrature points.  Similarly, let $\tilde{\bm{V}_f}{P}_f$ denote the matrix which projects to the trace space using full-side quadrature and evaluates at split-side quadrature points
\begin{align*}
\bm{C}^i_f &= \begin{pmatrix}
\bm{0} & {\rm diag}\LRp{\hat{\bm{n}}_i}{\bm{V}_f}\tilde{\bm{P}}_f\\
-\tilde{\bm{V}_f}{\bm{P}}_f{\rm diag}\LRp{\hat{\bm{n}}_i} &\bm{0}
\end{pmatrix}, \\
\bm{S}^i_f &= \begin{pmatrix}
\diag{\bm{w}_f} & \\
& \diag{\tilde{\bm{w}}_f}
\end{pmatrix}\bm{C}^i_f.
\end{align*}
It is straightforward to show that the matrix $\bm{S}_f$ is skew-symmetric.

%{fig:gqcon_noncon}
Let $\tilde{\bm{F}}_S$ denote the flux matrix, then 
\[
\begin{bmatrix}\bm{L}_f & \tilde{\bm{L}}_f\end{bmatrix} \LRp{\bm{C}^i_f\circ\tilde{\bm{F}}_S}\bm{1}
\]
is a skew-symmetric correction term which ensures entropy stability and high order accuracy (through consistency up to quadrature order).

Since the difference is high order accurate, you can remove this correction term and still get high order accuracy.  However, the spatial entropy RHS is no longer zero if you use inconsistent surface and coupling quadrature without the correction term (\note{prove this}).

\note{Do we need 3 layers of surface nodes in 3D?  Interpolation to isotropic split-side nodes is a Kronecker product - does that reduce things down?  }
